---
title: 'Multilevel linear models: varying slopes, non-nested models, and other complexities'
author:
date:
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

```{r arm13-pre, echo = F, message = F}
knitr::opts_chunk$set(
  cache = T,
  cache.path = '../arm_cache/',
  fig.path = '../arm_fig/',
  message = F,
  warning = F
  )
library(arm)
library(hrbrthemes)
library(stargazer)
load_tidy()

# arm link
arm_url <- 'http://stat.columbia.edu/~gelman/arm/examples/'

# arm functions
walk(list.files('./arm/arm_fun', full.names = T), source)
```

## Question 1

```{r arm13-q01_}

# reload chapter 3, exercise 5
dat <- knitr::load_cache('arm03-q05_', 'dat')

```

#### (a)

The model is given by:

$$
\begin{align}
y_i &\sim \text{N} \left( \alpha_{j[i]} + \beta_{j[i]} x_i + Z_i \Pi, \sigma_y^2 \right) \\
\begin{matrix}
  \alpha_j \\
  \beta_j
\end{matrix} &\sim \text{N}
\left(
  \begin{matrix}
  \mu_{\alpha} \\
  \mu_{\beta}
  \end{matrix},
  \begin{pmatrix}
  \sigma_{\alpha}^2 & \rho \sigma_{\alpha} \sigma_{\beta} \\
  \rho \sigma_{\alpha} \sigma_{\beta} & \sigma_{\beta}^2
  \end{pmatrix}
\right)
\end{align}
$$

where $x_i$ is a measure of the professor's beauty and $Z_i$ contains indicators for whether the professor is female and whether the professor is a minority.  In this case, the grouping variable course is created using the class* indicators; data with no indicator were dropped.

### (b)

```{r arm13-q01b}

# rename outcome variable and create id
dat %<>%
  rename(ce = courseevaluation) %>%
  mutate(
    id = 1:n()
  )

# create course category?
category <- select(dat, id, starts_with("class"))
category %<>% gather(course, value, -id)
category %<>% filter(value != 0)
category$course %<>% str_remove_all("class")
category$course %<>% as.numeric()

# join back category
dat %<>% left_join(select(category, id, course), by = "id")

# drop course = NA
dat %<>% filter(!is.na(course))

# fit the model
reg <- lmer(
  ce ~ btystdave + female + minority + (1 + btystdave | course),
  data = dat
  )
display(reg)

```

In this model, the unexplained within-course variation has an estimated standard deviation of $\sigma_{ce} = 0.43$, the estimated standard deviation of the course intercepts is $\sigma_{course} = 0.33$ and the estimated standard deviation of the course slopes is $\sigma_{bystdave} = 0.17$.

### (c)

```{r arm13-q01c-1}

# coefficients
dat_coef <- coef(reg)$course %>%
  rownames_to_column(var = "course") %>%
  mutate_at(vars(course), as.numeric)

# create new grouping
dat %<>%
  mutate(
    female_minority = case_when(
      female == 0 & minority == 0 ~ "White male",
      female == 0 & minority == 1 ~ "Minority male",
      female == 1 & minority == 0 ~ "White female",
      TRUE ~ "Minority female"
    )
  )
  
# plot
p <- ggplot() +
  geom_point(
    aes(x = btystdave, y = ce, colour = female_minority),
    data = dat
  ) +
  geom_abline(
    aes(intercept = `(Intercept)`, slope = btystdave),
    data = dat_coef
  ) +
  scale_colour_ipsum() +
  theme_ipsum() +
  labs(x = "Beauty score", y = "Evaluation score", colour = "Type") +
  facet_wrap(~course)
p

```

```{r arm13-q01c-2}

pdat <- bind_cols(
  coef(reg)$course[, 1], 
  ranef(reg)$course[, 1],
  .name_repair = "unique"
)
names(pdat) <- c("est", "se")

p <- ggplot(pdat, aes(x = factor(1:30), colour = factor(1:30))) +
  geom_point(aes(y = est)) +
  geom_errorbar(aes(ymin = est - 2 * se, ymax = est + 2 * se)) +
  theme_ipsum() +
  labs(x = "Course", y = "Intercept") +
  guides(colour = F)
p

```

```{r arm13-q01c-3}

pdat <- bind_cols(
  coef(reg)$course[, 2], 
  ranef(reg)$course[, 2],
  .name_repair = "unique"
)
names(pdat) <- c("est", "se")

p <- ggplot(pdat, aes(x = factor(1:30), colour = factor(1:30))) +
  geom_point(aes(y = est)) +
  geom_errorbar(aes(ymin = est - 2 * se, ymax = est + 2 * se)) +
  theme_ipsum() +
  labs(x = "Course", y = "Slope") +
  guides(colour = F)
p

```

## Question 2

### (a)

### Question 3

```{r arm13-q03_}

# reload chapter 3, exercise 5
dat <- knitr::load_cache('arm11-q03_', 'dat')

```

### (a)
