---
title: 'ARM: Chapter 9'
output: html_document
---

```{r arm09-pre, echo = F, message = F}
knitr::opts_chunk$set(
  cache = T,
  cache.path = 'arm_cache/',
  fig.path = 'arm_fig/',
  message = F,
  warning = F
  )
library(arm)
load_tidy()

# arm link
arm_url <- 'http://stat.columbia.edu/~gelman/arm/examples/'

# arm functions
walk(list.files('./arm/arm_fun', full.names = T), source)
```

### Question 1

Randomly assign vending machine availability at each school.

### Question 2

Obviously, you cannot randomly assign people to smoke.  One possibily might be to randomly assign cigarette taxes; as this would incentive some smokers to quit you could plausibly measure the effect the tax had on lung cancer.

### Question 3

Skip

### Question 4

Create the data:

```{r arm09-q04_}
dat <- tibble(
  category = 1:8,
  persons = c(rep(300, 2), rep(500, 2), rep(200, 4)),
  x  = factor(rep(c(0, 1), 4)),
  Tr = factor(rep(c(0, 0, 1, 1), 2)),
  y0 = c(rep(4, 4), rep(10, 4)),
  y1 = c(rep(6, 4), rep(12, 4))
)
dat %<>% mutate(y = if_else(Tr == 1, y1, y0))
```

#### (a)

The treatment effect is two for all categories so the average is two.

#### (b)

```{r arm09-q04b}
p <- dat %>%
  unite(id, x, y0, y1, remove = F) %>%
  ggplot() +
  geom_bar(
    aes(id, persons, fill = Tr),
    stat = 'identity',
    position = 'dodge'
    )
p
```

More of the individuals with potential outcomes $(y_0, y_1) = (4, 6)$ received treatment than did not suggesting treatment assignment was not random.

#### (c)

```{r arm09-q04c}
dat %>%
  group_by(Tr) %>%
  summarise(y = weighted.mean(y, persons)) %>%
  use_series('y') %>%
  diff()
```

This value is less than the true average treatment effect.  Given (c), this would be expected.

#### (d)

```{r arm09-q04d}
dat %>%
  group_by(x) %>%
  count(y0, wt = persons)
dat %>%
  group_by(x) %>%
  count(Tr, wt = persons)
```

$x$ does not appear to be a confounder here.  The potential outcomes are the same regardless of level and the probability of treatment is also the same.

### Question 5

```{r arm09-q05_}
reg <- lm(y ~ Tr + x, data = dat, weights = persons)
summary(reg)
```

### Question 6

This would be ill advised.  While access to therapy has been randomized, the number of sessions attended has not.  Among other potential issues, there may be selection bias stemming from the decision of how many sessions to actually attend.

### Question 7

The original regressions is:
$$
y_i = \alpha + \theta T_i + \beta x_i + \text{error}_i
$$
while the gain-score model is given by:
$$
g_i = \delta + \gamma T_i + \zeta x_i + \text{error}_i
$$
where $g_i = y_i - \bar x$.  Substituting in for $g_i$ and rearranging implies:
$$
y_i = (\delta + \bar x) + \gamma T_i + \zeta x_i + \text{error}_i
$$
which implies:
$$
\begin{align}
\alpha &= \delta + \bar x \\
\beta &= \zeta \\
\theta &= \gamma
\end{align}
$$

### Question 8

First, generate some data:

```{r arm09-q08-1}
set.seed(1)
dat <- tibble(
  x  = runif(100, 0, 100),
  y0 = x + rnorm(100, 0, 5),
  Tr = factor(rbinom(100, 1, 0.5)),
  a = y0,
  b = y0 + 20,
  c = y0 + 0.5 * x
)
dat %<>% gather(scenario, y1, -1:-3)
dat %<>% mutate(y = if_else(Tr == 1, y1, y0))
```

Next, estimate the models:

```{r arm09-q08-2}
reg_fun <- function(model) {
  lm(y ~ Tr + x + Tr:x, data = filter(dat, scenario == model))
}
reg <- map(c('a', 'b', 'c'), reg_fun) ; names(reg) <- c('a', 'b', 'c')
reg_coef <- map(reg, coef) ; names(reg_coef) <- c('a', 'b', 'c')
```

Finally, plot:

```{r arm09-q08-3}
p <- ggplot(dat) +
  geom_point(aes(x, y, shape = Tr)) +
  scale_shape_manual(values = c(16, 1)) +
  facet_wrap(~scenario) +
  geom_abline(
    aes(intercept = reg_coef$a[1], slope = reg_coef$a[3]),
    col = 'red',
    data = filter(dat, scenario == 'a')
  ) +
  geom_abline(
    aes(intercept = sum(reg_coef$a[1:2]), slope = sum(reg_coef$a[3:4])),
    col = 'red', linetype = 2,
    data = filter(dat, scenario == 'a')
  ) +
  geom_abline(
    aes(intercept = reg_coef$b[1], slope = reg_coef$b[3]),
    col = 'blue',
    data = filter(dat, scenario == 'b')
  ) +
  geom_abline(
    aes(intercept = sum(reg_coef$b[1:2]), slope = sum(reg_coef$b[3:4])),
    col = 'blue', linetype = 2,
    data = filter(dat, scenario == 'b')
  ) +
  geom_abline(
    aes(intercept = reg_coef$c[1], slope = reg_coef$c[3]),
    col = 'green',
    data = filter(dat, scenario == 'c')
  ) +
  geom_abline(
    aes(intercept = sum(reg_coef$c[1:2]), slope = sum(reg_coef$c[3:4])),
    col = 'green', linetype = 2,
    data = filter(dat, scenario == 'c')
  )
p
```

### Question 9

First, generate some data:

```{r arm09-q09_}
set.seed(1)
dat <- tibble(
  x  = runif(100, 0, 100),
  y0 = x + rnorm(100, 0, 5),
  Tr = factor(rbinom(100, 1, 0.5)),
  a = y0 + 20,
  b = y0 + 0.5 * x ^ 2
)
dat %<>% gather(scenario, y1, -1:-3)
dat %<>% mutate(y = if_else(Tr == 1, y1, y0))
```

#### (a)

```{r arm09-q09a}
dat %>%
  filter(scenario == 'a') %>%
  group_by(Tr) %>%
  summarise_at(vars(y), mean) %>%
  use_series('y') %>%
  diff()
lm(y ~ Tr + x, data = filter(dat, scenario == 'a'))$coefficients[[2]]
```

#### (b)

```{r arm09-q09b}
reg <- list(
  r1 = lm(y ~ Tr + x + x:Tr, data = filter(dat, scenario == 'b')),
  r2 = lm(y ~ Tr + x + I(x ^ 2):Tr, data = filter(dat, scenario == 'b'))
)
reg_coef <- map(reg, coef) ; names(reg_coef) <- c('r1', 'r2')

plot_fun <- function(x, coef = NULL) {
  coef$a + coef$b * x + coef$c * x ^ 2
}
c1 <- list(a = reg_coef$r2[1], b = reg_coef$r2[3], c = reg_coef$r2[4])
c2 <- list(a = sum(reg_coef$r2[1:2]), b = reg_coef$r2[3], c = reg_coef$r2[5])

p <- ggplot(filter(dat, scenario == 'b')) +
  geom_point(aes(x, y, shape = Tr)) +
  scale_shape_manual(values = c(16, 1)) +
  geom_abline(
    aes(intercept = reg_coef$r1[1], slope = reg_coef$r1[3]),
    col = 'red'
  ) +
  geom_abline(
    aes(intercept = sum(reg_coef$r1[1:2]), slope = sum(reg_coef$r1[3:4])),
    col = 'red', linetype = 2
  ) +
  stat_function(
    fun = plot_fun, args = list(coef = c1),
    col = 'blue'
  ) +
  stat_function(
    fun = plot_fun, args = list(coef = c2),
    col = 'blue', linetype = 2
  )
p
```

### Question 10

Load the data:

```{r arm09-q10_}
dat <- read_stata(str_c(arm_url, 'sesame/sesame.dta'))
dat %<>% mutate(Tr = factor(if_else(viewenc == 1 | encour == 1, 1, 0)))
```

#### (a)

Practically speaking, it might be difficult to get children to actually watch the program; encouraging them to do so allows them to accept the information on their own terms.  Actual watching would also likely be more difficult to randomize since it would require some level of commitment from the subjects.

#### (b)

This would require the assumption that test scores would not change for those children who were not encouraged to watch the show.

```{r arm09-q10b}
tests <- c('body', 'classify', 'forms', 'letters', 'numbers', 'relations')

dat %<>%
  select(Tr, contains('pre'), contains('post')) %>%
  mutate(id = row_number()) %>%
  gather(variable, value, -id, -Tr) %>%
  extract(variable, c('time', 'test'), '(^pre|^post)([^pre]+|[^post]+)') %>%
  spread(time, value)
dat$test %<>% factor(labels = tests)

p <- ggplot(aes(pre, post), data = dat) +
  geom_point(aes(shape =  Tr)) + 
  geom_smooth(
    method = 'lm', se = F,
    col = 'red', linetype = 2,
    data = filter(dat, Tr == 1)
    ) +
  geom_smooth(
    method = 'lm', se = F,
    col = 'red', 
    data = filter(dat, Tr == 0)
    ) +
  scale_shape_manual(values = c(16, 1)) +
  facet_wrap(~test, scales = 'free')
p
```

Given the plots above, this does not seem realistic; test scores improve for both the control units (dots and solid lines) and the treated units (circles and dashed lines).

### Question 11

Load the data:

```{r arm09-q11_}
dat <- knitr::load_cache('arm09-q10_', 'dat')
```

#### (a)

```{r arm09-q11a}
reg1 <- lm(postlet ~ Tr + prelet, data = dat)
summary(reg1)

reg2 <- lm(postnumb ~ Tr + prenumb, data = dat)
summary(reg2)
```

In both cases, treatment had a positive effect but it is only statistically significant for the test on letters.

#### (b)

```{r arm09-q11b}
reg1 <- lm(postlet ~ Tr:factor(regular) + prelet - 1, data = dat)
summary(reg1)
c1 <- summary(reg1)$coefficients[, 1]
s1 <- summary(reg1)$coefficients[, 2]
1 - pt((c1[[5]] - c1[[4]]) / sqrt(s1[[5]] ^ 2 + s1[[4]] ^ 2), 231)

reg2 <- lm(postnumb ~ Tr:factor(regular) + prenumb - 1, data = dat)
summary(reg2)
c2 <- summary(reg2)$coefficients[, 1]
s2 <- summary(reg2)$coefficients[, 2]
1 - pt((c2[[5]] - c2[[4]]) / sqrt(s2[[5]] ^ 2 + s2[[4]] ^ 2), 231)
```

After interacting the treatment variable with frequency of viewing, treatment has only a small positive effect for those who watched the show regularly; these effects were not statistically significant. 

#### (c)

Given the description of the data in Question 10, it seems more plausible that the estimates in (a) have a casual interpretation.  Frequency of viewing would seem likely to introduce selection bias.

### Question 12

Load the data:

```{r arm09-q12_}
dat <- read.table(str_c(arm_url, 'cows/cow.dat'), skip = 4, header = T)
```

#### (a)

```{r arm09-q12a}
rega <- lm(fat ~ level, data = dat)
summary(rega)
```

The regression does not include the pre-treatment predictors: lactation, age, and initial weight.  As treatment assignment was based on this information it should be controlled for.

#### (b)

```{r arm09-q12b}
regb <- lm(
  fat ~ level + lactation + age + initial.weight,
  data = dat
  )
summary(regb)
```

Here, each of the pre-treatment predictors are included in the model.

#### (c)

```{r arm09-q12c}
regc <- lm(
  fat ~ factor(level) + lactation + age + initial.weight,
  data = dat
  )
summary(regc)

pdat <- tidy(regc)
pdat <- pdat[1:4, ]
pdat$estimate[2:4] <- pdat$estimate[2:4] + pdat$estimate[1]
pdat %<>% 
  mutate(
    lower = estimate - 2 * std.error,
    upper = estimate + 2 * std.error
  )

p <- ggplot(pdat, aes(term, estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_hline(aes(yintercept = regb$coefficients[2]), col = 'red') +
  scale_x_discrete(labels = str_c('Level = ', seq(0, 0.3, by = 0.1)))
p
```

### Question 13

#### (a)

```{r arm09-q13a}
dat88 <- read.table(str_c(arm_url, 'congress/cong3/1988.asc'))
names(dat88) <- c('icpsr', 'district', 'incumb', 'dem', 'rep')
dat88 %<>% 
  mutate(
    dshare = dem / (dem + rep),
    iparty = if_else(dshare > 0.5, 1, -1) 
    )

dat90 <- read.table(str_c(arm_url, 'congress/cong3/1990.asc'))
names(dat90) <- c('icpsr', 'district', 'incumb', 'dem', 'rep')
dat90 %<>% mutate(dshare = dem / (dem + rep))

dat <- left_join(dat90, dat88, by = c('icpsr', 'district'))
dat %<>% filter(dshare.x > 0.1 & dshare.x < 0.9)
dat %<>% filter(dshare.y > 0.1 & dshare.y < 0.9)

reg <- lm(dshare.x ~ dshare.y + incumb.x + iparty, data = dat)
summary(reg)
```

#### (b)

```{r arm09-q13b}
x <- coef(reg)
y1 <- x[1] + x[2] * c(0.0, 0.5) - x[3] - x[4]
y2 <- x[1] + x[2] * c(0.5, 1.0) + x[3] + x[4]

p <- ggplot(dat) +
  geom_point(aes(dshare.y, dshare.x, col = factor(incumb.x)), alpha = 1) +
  scale_color_manual(
    name = 'incumbency',
    labels = c('rep', 'none', 'dem'),
    values = c('red', 'green', 'blue')
    ) +
  geom_segment(aes(x = 0.0, xend = 0.5, y = y1[1], yend = y1[2])) +
  geom_segment(aes(x = 0.5, xend = 1.0, y = y2[1], yend = y2[2])) +
  labs(x = 'dem share 1988', y = 'dem share 1990')
p
```

#### (c)

The casual effect of incumbency is valid for the data where $v_{i,t-2} \approx 0.5$; winners of close elections (and thus incumbents) are essentially randmonly assigned in these cases.

### Question 14

Skip