---
title: 'ARM: Chapter 4'
output: html_document
---

```{r arm04-pre, echo = F, message = F}
knitr::opts_chunk$set(
  cache = T,
  cache.path = 'arm_cache/',
  fig.path = 'arm_fig/',
  message = F,
  warning = F
  )
library(arm)
load_tidy()

# arm link
arm_url <- 'http://stat.columbia.edu/~gelman/arm/examples/'
```

### Question 1

#### (a)

The factors are $e^{-0.25}$ and $e^{0.25}$.

#### (b)

Import the height data used in Chapter 4, simulate weights, and plot:

```{r arm04-q01b}
set.seed(1)

dat <- read_stata(str_c(arm_url, 'earnings/heights.dta'))

dat$weight <- exp(-3.5 + 2 * log(dat$height) + rnorm(nrow(dat), 0, 0.25))

p <- ggplot(dat) +
  geom_point(aes(log(height), log(weight))) +
  geom_abline(aes(intercept = -3.5, slope = 2), col = 'red')
p
```

### Question 2

#### (a)

Remove missing values from the data:

```{r arm04-q02a}
dat <- na.omit(dat)
```

#### (b)

Center height:

```{r arm04-q02b}
dat$height_c <- dat$height - mean(dat$height, na.rm = T)

reg <- lm(earn ~ height_c, data = dat) ; summary(reg)
```

So a person of average height earns \$`r format(reg$coefficients[1], nsmall = 2, big.mark = ',')` on average.

#### (c)

Skip

### Question 3

Create similar dataset and models:

```{r arm04-q03_}

# age (the original paper indicates the survey was conducted in 1990)
dat$age <- 90 - dat$yearbn
dat$age[dat$age <= 0] <- 1990 - (1800 + dat$yearbn[dat$age <= 0])

# age bins
age_brks <- c(0, 18, 29, 44, 64, Inf)
dat <- dat %>%
  mutate(
    age_10  = age / 10,
    age_bin = cut(age, breaks = age_brks, include.lowest = T)
  )

# models
reg1 <- lm(weight ~ age_10, data = dat)
reg2 <- lm(weight ~ age_10 + I(age_10 ^ 2), data = dat)
reg3 <- lm(weight ~ age_bin, data = dat)
```

#### (a) to (b)

```{r arm04-q03a}
reg2_fun <- function(x) (-0.0157) / 100 * x ^ 2 + (-2.13) / 10 * x + 148

p <- ggplot(dat) +
  geom_point(aes(age, weight)) +
  geom_abline(
    aes(intercept = reg1$coefficients[1], slope = reg1$coefficients[2] / 10),
    col = 'red'
  ) +
  stat_function(fun = reg2_fun, col = 'blue')
p
```

#### (c)

```{r arm04-q03b}
p <- ggplot(filter(dat, age_bin != '[0,18]')) + 
  geom_point(
    aes(age, weight),
    col = 'grey',
    data = select(filter(dat, age_bin != '[0,18]'), age, weight)
  ) +
  geom_point(aes(age, weight, colour = age_bin)) +
  scale_colour_manual(values = c("yellow", "green", "blue", "red")) +
  geom_hline(
    aes(yintercept = reg3$coefficients[1]),
    col = 'yellow',
    data = filter(dat, age_bin == '(18,29]')
  ) +
  geom_hline(
    aes(yintercept = reg3$coefficients[1] + reg3$coefficients[2]),
    col = 'green',
    data = filter(dat, age_bin == '(29,44]')
  ) +
  geom_hline(
    aes(yintercept = reg3$coefficients[1] + reg3$coefficients[3]),
    col = 'blue',
    data = filter(dat, age_bin == '(44,64]')
  ) +
  geom_hline(
    aes(yintercept = reg3$coefficients[1] + reg3$coefficients[4]),
    col = 'red',
    data = filter(dat, age_bin == '(64,Inf]')
  ) +
  facet_grid(~age_bin) +
  theme(legend.position = 'none')
p
```

### Question 4

Import the data:

```{r arm04-q04_}
dat <- read_stata(str_c(arm_url, 'pollution/pollution.dta'))
```

#### (a)

```{r arm04-q04a}
p1 <- ggplot(dat) +
  geom_point(aes(nox, mort))
p1

reg <- lm(mort ~ nox, data = dat) ; summary(reg)

dat <- dat %>%
  mutate(
    pred = predict(reg, data = dat),
    resid = mort - pred
  )

p2 <- ggplot(dat) +
  geom_point(aes(pred, resid))
p2
```

Linearity seems like a stretch here.  The outlier where nox is 319 results in a poor fit.  The residual plot confirms this approach is problematic.

#### (b)

Let's try using the log of nox:

```{r arm04-q04b}
p1 <- ggplot(dat) +
  geom_point(aes(log(nox), mort))
p1

reg <- lm(mort ~ log(nox), dat = dat) ; summary(reg)

dat <- dat %>%
  mutate(
    pred = predict(reg, data = dat),
    resid = mort - pred
  )

p2 <- ggplot(dat) +
  geom_point(aes(pred, resid))
p2

```

This works better.

#### (c)

A one percent increase in nox is associated with an increase of mortality of `r format(reg$coefficients[2], nsmall = 1)`.

#### (d)

```{r arm04-q04d}
dat_long <- gather(dat, poll, value, hc:so2)

p1 <- ggplot(dat_long) +
  geom_point(aes(value, mort)) +
  facet_grid(~poll)
p1

reg <- lm(mort ~ log(hc) + log(nox) + log(so2), data = dat) ; summary(reg)

c1 <- reg$coefficients
c2 <- apply(dat[c('hc' ,'nox', 'so2')], 2, mean) %>% log()
c3 <- c1[2:4] * c2  

p2 <- ggplot(dat_long) +
  geom_point(aes(log(value), mort)) +
  facet_grid(~poll) +
  geom_abline(
    aes(intercept = c1[1] + c3[2] + c3[3], slope = c1[2]),
    col = 'red',
    data = filter(dat_long, poll == 'hc')
  ) +
  geom_abline(
    aes(intercept = c1[1] + c3[1] + c3[3], slope = c1[3]),
    col = 'red',
    data = filter(dat_long, poll == 'nox')
  ) +
  geom_abline(
    aes(intercept = c1[1] + c3[1] + c3[2], slope = c1[4]),
    col = 'red',
    data = filter(dat_long, poll == 'so2')
  )
p2
```

Given a single pollutant with the levels of the other two fixed (perhaps at their mean values), the coefficients correspond to the change in mortality for a one percent change in the given pollutant.

#### (e)

```{r arm04-q04e}
reg <- lm(mort ~ log(hc) + log(nox) + log(so2), data = dat[1:30, ]) ; summary(reg)

dat <- dat %>%
  mutate(
    pred = predict(reg, newdata = dat),
    resid = mort - pred,
    code = c(rep('train', 30), rep('test', 30))
  )

p <- ggplot(dat) +
  geom_point(aes(pred, mort, colour = code)) +
  geom_abline(intercept = 0, slope = 1)
p
```

### Question 5

#### (a)

* This measure implicitly assumes that a dollar has the same impact on vote share across districts.  This is likely unrealistic; for example, advertising costs are likely to vary by district.
* This measure improves on the last but may be problematic in uncontested districts; it also may be skewed by noncompetitive districts.
* This measure improves on the last in terms of skewness but still may be problematic in uncontested districts.
* This measure seems the best as it is standardized, ranging between 0 and 1, and has a clear demarcation at one half where the candidates raise equal amounts. 

#### (b)

One possiblity would be to use the last measure $M_i = D_i / (D_i + R_i)$ but introduce nonlinearity using polynomial transformations.

### Question 6

This represents an elasticity.  Most likely the authors meant $\beta = -0.3$; economists often consider elasticities in absolute terms.  This means that a one percent increase in price is associated with a 0.3 percent decrease in quantity.  As expected, demand for cigarettes is inelastic.

### Question 7

Skip

### Question 8

Skip
